### 2020-04-04 thesis/content/chapter-ltd.tex
completed 2-wl-gnn expressive power section

The convolution operator from \cref{defn:ltd:wl2-conv-step} is parameterized by the three matrices $W_{\mathrm{L}}^{(t)}, W_{\mathrm{F}}^{(t)},  W_{\Gamma}^{(t)} \in \mathbb{R}^{d^{(t-1)} \times d^{(t)}}$ and uses two freely choosable nonlinearities $\sigma$ and $\sigma_{\Gamma}$.

@@

		By choosing $W_{F}^{(t)} = \mathbf{0}$ the convolution behaves like a regular fully connected layer.

@@

		Instead of filtering out all neighbors, $W_{F}^{(t)}$ also allows to filter the neighborhood of $e_{ij}$ more selectively.

@@

	Let $\mathit{Pool}_2(Z_G^{(T,2+K)}) \coloneqq \mathit{Pool}_1(\ldblbrace z_{ij}\, |\, z_{ij} = \varphi(Z_G^{(T,2+K)}[e_{ij}]) \land e_{ij} \in \mathcal{E}_{G^1} \land z_{ij} \neq \texttt{nil} \rdblbrace)$.

@@

			${Z_G^{(t,2+K)}[e_{ij}]}_1 = \mathbbm{1}[i = j]$, i.e.\ the first component of each 2-\acs{wl} feature vector allows $\varphi$ to decide whether it should be mapped to \texttt{nil}.

@@

			${Z_G^{(t,2+K)}[e_{ii}]}_{2, \dots, (d^{(t)} + 1)} = Z_G^{(t)}[v_{i}]$, i.e.\ the second to $(1 + d^{(t)})$-th components of each self-loop feature vector contain the corresponding convolved vertex feature vector at layer $t$.

@@

	As previously mentioned, a 2-\acs{wl} convolution layer is reduced to a fully connected layer if $W_{F}^{(t)} = \mathbf{0}$.

@@

	Using invariants~\ref{inv:ltd:wl2-simulation:indicator} and~\ref{inv:ltd:wl2-simulation:feature} for $t = T$ we can therefore set
	\begin{align*}
		\varphi(Z_G^{(T,2+K)}[e_{ij}]) \coloneqq \begin{cases}
			{Z_G^{(T,2+K)}[e_{ij}]}_{2, \dots, (d^{T} + 1)} & \text{if ${Z_G^{(T,2+K)}[e_{ij}]}_1 = 1$} \\
			\texttt{nil} & \text{else}
		\end{cases}
		\text{.}
	\end{align*}
	By our previous definition of $\mathit{Pool}_2$, this in turn implies that $\mathit{Pool}_2(Z_G^{(T,2+K)}) = \mathit{Pool}_1(Z_G^{(T)}) \iff h_2(G) = h_1(G)$ which concludes the proof.
\end{proof}

>>>>

The convolution operator from \cref{defn:ltd:wl2-conv-step} is parameterized by the three matrices $W_{\mathrm{L}}^{(t)}, W_{\mathrm{F}}^{(t)},  W_{\Gamma}^{(t)} \in \mathbb{R}^{d^{(t-1)} \times d^{(t)}}$ and uses two freely choosable activation functions $\sigma$ and $\sigma_{\Gamma}$.

@@

		By choosing $W_{\mathrm{F}}^{(t)} = \mathbf{0}$ the convolution behaves like a regular fully connected layer.

@@

		Instead of filtering out all neighbors, $W_{\mathrm{F}}^{(t)}$ also allows to filter the neighborhood of $e_{ij}$ more selectively.

@@

	Let $\mathit{Pool}_2\left(Z_G^{(T,2+K)}\right) \coloneqq \mathit{Pool}_1(\ldblbrace z_{ij}\, |\, z_{ij} = \varphi\left(Z_G^{(T,2+K)}[e_{ij}]\right) \land e_{ij} \in \mathcal{E}_{G^1} \land z_{ij} \neq \texttt{nil} \rdblbrace)$.

@@

			${Z_G^{(t,2+K)}[e_{ij}]}_1 = \mathbbm{1}[i = j]$, i.e.\ the first component of each 2-\acs{wl} feature vector allows $\varphi$ to decide whether that vector should be mapped to \texttt{nil}.

@@

			${Z_G^{(t,2+K)}[e_{ii}]}_{2, \dots, (d^{(t)} + 1)} = Z_G^{(t)}[v_{i}]$, i.e.\ the second to $(1 + d^{(t)})$-th components of each self-loop feature vector in $h_2$ contain the corresponding convolved vertex feature vector at layer $t$ in $h_1$.

@@

	As previously mentioned, a 2-\acs{wl} convolution layer is reduced to a fully connected layer if $W_{\mathrm{F}}^{(t)} = \mathbf{0}$.

@@

	Using invariants~\ref{inv:ltd:wl2-simulation:indicator} and~\ref{inv:ltd:wl2-simulation:feature} for $t = T$ we can therefore set
	\begin{align*}
		\varphi\left(Z_G^{(T,2+K)}[e_{ij}]\right) \coloneqq \begin{cases}
			{Z_G^{(T,2+K)}[e_{ij}]}_{2, \dots, (d^{T} + 1)} & \text{if ${Z_G^{(T,2+K)}[e_{ij}]}_1 = 1$} \\
			\texttt{nil} & \text{else}
		\end{cases}
		\text{.}
	\end{align*}
	By our previous definition of $\mathit{Pool}_2$, this in turn implies that $\mathit{Pool}_2(Z_G^{(T,2+K)}) = \mathit{Pool}_1(Z_G^{(T)}) \iff h_2(G) = h_1(G)$ which concludes the proof.
\end{proof}
\begin{cor}\label{cor:ltd:wl2-gnn-wl1-power}
	2-\acs{wl}-\acsp{gnn} have at least the same discriminative power as 1-\acs{wl}.
\end{cor}
\begin{proof}
	The corollary directly follows from the fact that 2-\acs{wl}-\acsp{gnn} can simulate \acp{gin} by \cref{thm:ltd:wl2-simulation} and the fact that \acp{gin} have the same discriminative power as 1-\acs{wl} by \cref{prop:related:gcnn-wl1-limit} (see \cpageref{prop:related:gcnn-wl1-limit}) because they use injective vertex neighborhood hashing functions~\cite{Xu2018}.
\end{proof}

To complete our analysis of the expressive power of 2-\acs{wl}-\acsp{gnn} we now show that they are not just as powerful as 1-\acs{wl} but in fact more powerful than 1-\acs{wl}.
\begin{prop}\label{cor:ltd:wl2-gnn-regular}
	There are $d$-regular graphs $G$ and $H$ of size $n$ which can be distinguished by 2-\acs{wl}-\acsp{gnn}.
\end{prop}
\begin{proof}
	The proposition follows if we choose the six-cycle graph for $G$ and the two three-cycles graph for $H$, both of which should be well-known at this point.
	Let $h_2 = \mathit{Pool} \circ S$ be a 2-\acs{wl}-\acs{gnn} with the neighborhood radius $r = 1$ which consists of a single 2-\acs{wl} convolution layer $S: \mathbb{R}^{* \times 2} \to \mathbb{R}^{* \times 1}$ and the pooling layer $\mathit{Pool} = \min$.
	In accordance with \cref{defn:ltd:wl2-conv-init} we set the initial feature vectors of the vertices $v_i$ of $G$ and $H$ to $Z^{(0)}[e_{ii}] \coloneqq (1, 0)$ and the initial feature vectors of their edges $e_{ij}$ to $Z^{(0)}[e_{ij}] \coloneqq (0, 1)$.
	Let the weight matrices of $S$ be $W_{\mathrm{L}} \coloneqq \mathbf{0}$ and $W_{\mathrm{F}} = W_{\Gamma} \coloneqq \begin{pmatrix} 1\\1 \end{pmatrix}$.
	For simplicity we choose the identity activation functions $\sigma = \sigma_{\Gamma} = \mathrm{id}$.

	By \cref{defn:ltd:wl2-conv-step} all self-loops $e_{ii}$ of $G^1$ and $H^1$ have the three neighbors $\{ \textcolor{t_blue}{\ldblbrace e_{ii}, e_{ii} \rdblbrace},\allowbreak \textcolor{t_red}{\ldblbrace e_{ij}, e_{ji} \rdblbrace},\allowbreak \textcolor{t_darkgreen}{\ldblbrace e_{il}, e_{li} \rdblbrace} \}$, i.e.\ the length-two walk along $e_{ii}$ itself and the length-two walks to and from the two neighboring vertices $\Gamma(v_i) = \{ v_j, v_l \}$. % chktex 25
	Therefore the convolved feature vector of all self-loops are $Z^{(1)}[e_{ii}] = (1) \odot (\textcolor{t_blue}{(1 + 1)} + \textcolor{t_red}{(1 + 1)} + \textcolor{t_darkgreen}{(1 + 1)}) = 6$. % chktex 25
	However for the non-self-loops of $G^1$ and $H^1$, i.e.\ the edges of $G$ and $H$, we get differing convolved feature vectors.
	The 2-\acs{wl} neighbors of $e_{ij} \in \mathcal{E}_G$ are $\{ \ldblbrace e_{ii},  e_{ij} \rdblbrace,\allowbreak \ldblbrace e_{ij}, e_{jj} \rdblbrace \}$.
	The 2-\acs{wl} neighbors of $e_{ij}' \in \mathcal{E}_H$ are $\{ \ldblbrace e_{ii}',  e_{ij}' \rdblbrace,\allowbreak \ldblbrace e_{ij}',  e_{jj}' \rdblbrace,\allowbreak \ldblbrace e_{il}',  e_{lj}' \rdblbrace \}$ where $v_l' \in \mathcal{V}_H$ is the common neighbor of $v_i'$ and $v_j'$.
	The different neighborhood sizes of the edges of $G$ and $H$ imply that $\forall e_{ij} \in \mathcal{E}_G: Z^{(1)}[e_{ij}] = 4$ while $\forall e_{ij}' \in \mathcal{E}_H: Z^{(1)}[e_{ij}] = 6$.
	Thus $h_2(G) = \min\{ 4, 6 \} \neq \min\{ 6, 6 \} = h_2(H)$ which concludes the proof. % chktex 21
\end{proof}
\begin{cor}\label{cor:ltd:wl2-gnn-more-wl1-power}
	The discriminative power of 2-\acs{wl}-\acsp{gnn} is strictly greater than that of the 1-\acs{wl} algorithm.
\end{cor}
\begin{proof}
	The corollary directly follows from \cref{cor:ltd:wl2-gnn-wl1-power} and \cref{cor:ltd:wl2-gnn-regular} since 1-\acs{wl} cannot distinguish regular graphs~\cite[cor.~1.8.5]{Immerman1990}.
\end{proof}

This concludes our analysis of the expressive power of 2-\acs{wl}-\acsp{gnn}.
The key takeaway from this section is that 2-\acs{wl}-\acsp{gnn} are more powerful than all vertex neighborhood aggregation \acp{gcnn} because the power of the latter is upper bounded by 1-\acs{wl}.
Additionally we can conclude that 2-\acs{wl}-\acsp{gnn} are in fact also more powerful than 2-\acsp{gnn} due to \cref{prop:ltd:2gnn-regular-limit}.
A more thorough discussion of the relation between 2-\acs{wl}-\acsp{gnn} and 2-\acs{wl} is left for future work.

### 2020-04-04 thesis/content/chapter-related-work.tex
completed 2-wl-gnn expressive power section

\subsubsection{The $k$-dimensional \acs{wl} algorithm}
As we just saw, the 1-\acs{wl} algorithm iteratively refines colorings of single vertices.
While the obtained colorings differ for most non-isomorphic graphs $G \not\simeq H$, 1-\acs{wl} does not generally solve the \ac{gi} problem as illustrated in \cref{fig:related:wl1-problem}.
\begin{figure}[ht]
	\centering
	\includegraphics[width=\linewidth]{gfx/related-work/wl1-problem.pdf}
	\caption[Two simple non-isomorphic graphs that are indistinguishable by 1-\acs{wl}.]{
		Two simple non-isomorphic graphs that are indistinguishable by 1-\acs{wl}.
	}\label{fig:related:wl1-problem}
\end{figure}\\
By extending \ac{wl} to higher dimensions, such 1-\acs{wl} indistinguishable cases can however be handled.
Analogous to the 1-dimensional \cref{defn:related:wl1-refine}, the $k$-dimensional color refinement step is defined by
\begin{defn}\label{defn:related:wlk-refine}
	$\begin{aligned}[t]
		\chi_{G,k}^{(i+1)}(s) \coloneqq h\left(\chi_{G,k}^{(i)}(s), \ldblbrace (\chi_{G,k}^{(i)}(s[u/1]), \dots, \chi_{G,k}^{(i)}(s[u/k]))\, |\, u \in \mathcal{V}_G \rdblbrace\right)\\
		\text{with } s = (v_1, \dots, v_k) \in \mathcal{V}_G^k,\quad s[u/j] \coloneqq (v_1, \dots, v_{j-1}, u, v_{j+1}, \dots, v_k) \text{.}
	\end{aligned}$
\end{defn}
In 1-\acs{wl} a vertex color is refined by combining the colors of neighboring vertices.
In $k$-\acs{wl} the color of a $k$-tuple $s \in \mathcal{V}_G^k$ is refined by combining the colors of its neighborhood which is defined as the set of all $k$-tuples in which at most one vertex differs from $s$.
Note that each vertex $k$-tuple has one neighbor for each $u \in \mathcal{V}_G$, each of which is a $k$-tuple of vertex $k$-tuples.
This more abstract notion of neighborhood is illustrated in \cref{fig:related:wl-neighbors}.
\begin{figure}[ht]
	\centering
	\begin{subfigure}{0.32\textwidth}
		\centering
		\includegraphics[width=0.8\textwidth]{gfx/related-work/wl1-neighbors.pdf}
		\subcaption{$v_1$'s 1-WL neighbors}\label{fig:related:wl-neighbors:1}
	\end{subfigure}%
	\begin{subfigure}{0.32\textwidth}
		\centering
		\includegraphics[width=0.8\textwidth]{gfx/related-work/wl2-neighbors.pdf}
		\subcaption{$(v_1, v_2)$'s 2-WL neighbors}\label{fig:related:wl-neighbors:2}
	\end{subfigure}%
	\begin{subfigure}{0.36\textwidth}
		\centering
		\includegraphics[width=0.711\textwidth]{gfx/related-work/wl3-neighbors.pdf}
		\subcaption{$(v_1, v_2, v_3)$'s 3-WL neighbors}\label{fig:related:wl-neighbors:3}
	\end{subfigure}
	\caption[\ac{wl} neighborhoods for different values of $k$.]{
		Tuple neighborhoods for different values of $k$.
		The vertices highlighted in \textcolor{t_blue}{blue} form the root tuple $s \in \mathcal{V}_G^k$ whose neighbors are shown;
		for simplicity neighbors with $u \in s$ are left out (see~\cref{defn:related:wlk-refine}).
		Each neighbor is highlighted with a different color, except for 3-WL where the \textcolor{t_red}{red}, \textcolor{t_darkgreen}{green} and \textcolor{t_darkyellow}{yellow} triples actually form the single neighbor for $u = v_4$.
	}\label{fig:related:wl-neighbors}
\end{figure}
For $k = 2$ this means that each potential edge $(v, w) \in \mathcal{V}_G^2$ has all possible paths of length $2$ from $v$ to $w$ as its neighbors (see~\cref{fig:related:wl-neighbors:2}).
Also note that, even though $k$-\acs{wl} refines $k$-tuple colors, lower-dimensional structures still get their own colors since a tuple does not have to consist of distinct vertices, i.e.\ in $k$-\acs{wl} the color of a single vertex $v \in \mathcal{V}_G$ is described by $\chi^{*}_{G, k}(s)$ for $s = (v, \dots, v) \in \mathcal{V}_G^k$.

@@

If we increase the \acs{wl}-dimension to $k = 2$, the described 1-\acs{wl} restrictions no longer apply.
2-\acs{wl} is able to distinguish more than $1 - \mathcal{O}(\nicefrac{1}{n})$ of the regular $n$-vertex graphs~\cite[cor.~1.8.6]{Immerman1990} and can even count cycles:
\begin{prop}[see \citet{Fuerer2017} and \citet{Arvind2019} for the full proof]\label{prop:related:wl2-cycle-count}
	2-\acs{wl} is able to count $m$-cycles for all $m \leq 7$ but it cannot even detect them for $m > 7$.
\end{prop}
\begin{hproof}
	Only the idea behind triangle and 4-cycle counting is outlined here to give an intuition for why \cref{prop:related:wl2-cycle-count} holds.
	Note that the 2-\acs{wl} neighbors of each edge $(v, w) \in \mathcal{E}_G$ correspond to all possible paths $(v, u, w)$ of length 2, i.e.\ all possible triangles.
	By \cref{defn:related:wlk-refine} there must thus be a color subset $C^{\triangle}_j \subseteq \mathcal{C}$ representing that an edge is involved in exactly $j$ triangles after one refinement step.
	2-\acs{wl} then trivially counts triangles by setting $g(\mathit{dist}_\chi) = \frac{1}{3} \sum_j j \sum_{c \in C^{\triangle}_j} \mathit{dist}_\chi(c)$ to satisfy \cref{defn:related:wl-compute}.
	Analogously for 4-cycles, let $C^{\square}_j \subseteq \mathcal{C}$ be the colors indicating a non-edge $(v, w) \notin \mathcal{E}_G$ with $j$ common vertex neighbors $u \in \Gamma_G(v) \cap \Gamma_G(w)$.
	The number of 4-cycles is then determined by the colors of the diagonals through them via $g(\mathit{dist}_\chi) = \frac{1}{2} \sum_{j \geq 2} \binom{j}{2} \sum_{c \in C^{\square}_j} \mathit{dist}_\chi(c)$.
	Using a similar but more involved combinatorial argument requiring multiple color refinement steps, 5-, 6- \& 7-cycle counting can be shown.
\end{hproof}
As we just saw, 2-\acs{wl} is significantly more powerful than 1-\acs{wl}.
Though, by \cref{prop:related:wl-upper}, there are of course still 2-\acs{wl} indistinguishable graphs;
among others those are the strongly regular graphs $\mathit{srg}_{n, d, \lambda, \mu}(G)$.
\begin{defn}
	$\begin{aligned}[t]
		\mathit{srg}_{n, d, \lambda, \mu}(G) \iff\ & {|\mathcal{V}_G|} = n \quad\land\quad \forall v \in \mathcal{V}_G: {|\Gamma_G(v)|} = d \\
		&\land \forall (v, w) \in \mathcal{E}_G: {|\Gamma_G(v) \cap \Gamma_G(w)|} = \lambda \\
		&\land \forall (v, w) \in \mathcal{V}_G^2 \setminus \mathcal{E}_G: {|\Gamma_G(v) \cap \Gamma_G(w)|} = \mu
	\end{aligned}$
\end{defn}
Generally this restriction of 2-\acs{wl} is not an issue since strongly regular graphs do not appear in typical \ac{gcr} datasets.
By going to $k = 3$, even some strongly regular graphs as well as all planar graphs can be distinguished though~\cite{Kiefer2017}.
Apart from that there are currently few results regarding the classes of distinguishable graphs and computable functions for even higher \ac{wl}-dimensions.

@@

\paragraph{A 1-\acs{wl} bound on \ac{gnn} power}
In the next step we will take a closer look at the relation between \acp{gcnn} that use vertex neighborhood convolution and 1-\acs{wl} by comparing their respective discriminative power.
\begin{prop}[see {\citet[lemma 2]{Xu2018}} for the proof]\label{prop:related:gcnn-wl1-limit}
	The discriminative power of a \ac{gcnn} that convolves vertex feature vectors via a convolution operator of the form $Z_G^{(t)}[v] = h^{(t)}\left(Z_G^{(t-1)}[v], \ldblbrace Z_G^{(t-1)}[u]\, |\, u \in \Gamma_G(v) \rdblbrace\right)$ is upper-bounded by that of 1-\acs{wl}, where $h^{(t)}: {(\mathbb{R}^{d^{(t-1)}})}^{*} \to \mathbb{R}^{d^{(t)}}$ is an arbitrary vertex neighborhood hashing function.
	Moreover, if $h^{(t)}$ is not injective, the \ac{gcnn} has a strictly lower discriminative power than 1-\acs{wl}.
\end{prop}
By \cref{prop:related:gcnn-wl1-limit} the \ac{gcn} architecture is strictly less powerful than 1-\acs{wl} due to the fact that $h_{\mathrm{\acs*{gcn}}}$ is not an injective hash function.
To overcome this limitation \citet{Xu2018} proposed an alternative vertex neighborhood convolution architecture called \ac{gin}:
\begin{align}
	Z_G^{(t)}[v] \coloneqq \mathrm{\acs*{mlp}}^{(t)}\left( Z_G^{(t-1)}[v] + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} Z_G^{(t-1)}[u] \right) \label{eq:related:gin-layer}
\end{align}
By leaving out the normalization factors $\mu(v)$ and by using a \ac{mlp} instead of the single fully-connected layer $\sigma \circ \Theta^{(t)}$ the resulting hashing function $h_{\mathrm{\acs*{gin}}}(z_1, \dots, z_m) = \mathrm{\acs*{mlp}}^{(t)}(\sum_{j=1}^{m} z_j)$ becomes injective\footnote{
	Under the assumption that the set of possible input feature matrices $X \in \mathbb{R}^{n \times d^{(0)}}$ is countable.
	This is reasonable since real-world \ac{gcr} domains typically only have a finite set of possible vertex feature vectors, e.g\ one-hot label encodings.
}, i.e.\ it assigns a unique color vector to each multiset of color vectors.
By \cref{prop:related:gcnn-wl1-limit} this implies that \ac{gin} has the same discriminative power as 1-\ac{wl}.
As expected, \ac{gin} therefore fits training data better than non-injective \acp{gcnn} like \ac{gcn}.
Interestingly \ac{gin} additionally seems to generalize better on test data than non-injective methods~\cite{Xu2018}\cite{Errica2020}.
It is not yet fully understood why this is the case.

>>>>

\subsubsection{The $k$-dimensional \acs{wl} algorithm}
As we just saw, the 1-\acs{wl} algorithm iteratively refines colorings of single vertices.
While the obtained colorings differ for most non-isomorphic graphs $G \not\simeq H$, 1-\acs{wl} does not generally solve the \ac{gi} problem as illustrated in \cref{fig:related:wl1-problem}.
\begin{figure}[ht]
	\centering
	\includegraphics[width=\linewidth]{gfx/related-work/wl1-problem.pdf}
	\caption[Two simple non-isomorphic graphs that are indistinguishable by 1-\acs{wl}.]{
		Two simple non-isomorphic graphs that are indistinguishable by 1-\acs{wl}.
	}\label{fig:related:wl1-problem}
\end{figure}\\
By extending \ac{wl} to higher dimensions, such 1-\acs{wl} indistinguishable cases can however be handled.
Analogous to the 1-dimensional \cref{defn:related:wl1-refine}, the $k$-dimensional color refinement step is defined by
\begin{defn}\label{defn:related:wlk-refine}
	$\begin{aligned}[t]
		\chi_{G,k}^{(i+1)}(s) \coloneqq h\left(\chi_{G,k}^{(i)}(s), \ldblbrace (\chi_{G,k}^{(i)}(s[u/1]), \dots, \chi_{G,k}^{(i)}(s[u/k]))\, |\, u \in \mathcal{V}_G \rdblbrace\right)\\
		\text{with } s = (v_1, \dots, v_k) \in \mathcal{V}_G^k,\quad s[u/j] \coloneqq (v_1, \dots, v_{j-1}, u, v_{j+1}, \dots, v_k) \text{.}
	\end{aligned}$
\end{defn}
In 1-\acs{wl} a vertex color is refined by combining the colors of neighboring vertices.
In $k$-\acs{wl} the color of a $k$-tuple $s \in \mathcal{V}_G^k$ is refined by combining the colors of its neighborhood which is defined as the set of all $k$-tuples in which at most one vertex differs from $s$.
Note that each vertex $k$-tuple has one neighbor for each $u \in \mathcal{V}_G$, each of which is a $k$-tuple of vertex $k$-tuples.
This more abstract notion of neighborhood is illustrated in \cref{fig:related:wl-neighbors}.
\begin{figure}[ht]
	\centering
	\begin{subfigure}{0.32\textwidth}
		\centering
		\includegraphics[width=0.8\textwidth]{gfx/related-work/wl1-neighbors.pdf}
		\subcaption{$v_1$'s 1-WL neighbors}\label{fig:related:wl-neighbors:1}
	\end{subfigure}%
	\begin{subfigure}{0.32\textwidth}
		\centering
		\includegraphics[width=0.8\textwidth]{gfx/related-work/wl2-neighbors.pdf}
		\subcaption{$(v_1, v_2)$'s 2-WL neighbors}\label{fig:related:wl-neighbors:2}
	\end{subfigure}%
	\begin{subfigure}{0.36\textwidth}
		\centering
		\includegraphics[width=0.711\textwidth]{gfx/related-work/wl3-neighbors.pdf}
		\subcaption{$(v_1, v_2, v_3)$'s 3-WL neighbors}\label{fig:related:wl-neighbors:3}
	\end{subfigure}
	\caption[\ac{wl} neighborhoods for different values of $k$.]{
		Tuple neighborhoods for different values of $k$.
		The vertices highlighted in \textcolor{t_blue}{blue} form the root tuple $s \in \mathcal{V}_G^k$ whose neighbors are shown;
		for simplicity neighbors with $u \in s$ are left out (see~\cref{defn:related:wlk-refine}).
		Each neighbor is highlighted with a different color, except for 3-WL where the \textcolor{t_red}{red}, \textcolor{t_darkgreen}{green} and \textcolor{t_darkyellow}{yellow} triples actually form the single neighbor for $u = v_4$.
	}\label{fig:related:wl-neighbors}
\end{figure}
For $k = 2$ this means that each potential edge $(v, w) \in \mathcal{V}_G^2$ has all possible walks of length $2$ from $v$ to $w$ as its neighbors (see~\cref{fig:related:wl-neighbors:2}).
Also note that, even though $k$-\acs{wl} refines $k$-tuple colors, lower-dimensional structures still get their own colors since a tuple does not have to consist of distinct vertices, i.e.\ in $k$-\acs{wl} the color of a single vertex $v \in \mathcal{V}_G$ is described by $\chi^{*}_{G, k}(s)$ for $s = (v, \dots, v) \in \mathcal{V}_G^k$.

@@

If we increase the \acs{wl}-dimension to $k = 2$, the described 1-\acs{wl} restrictions no longer apply.
2-\acs{wl} is able to distinguish more than $1 - \mathcal{O}(\nicefrac{1}{n})$ of the regular $n$-vertex graphs~\cite[cor.~1.8.6]{Immerman1990} and can even count cycles:
\begin{prop}[see \citet{Fuerer2017} and \citet{Arvind2019} for the full proof]\label{prop:related:wl2-cycle-count}
	2-\acs{wl} is able to count $m$-cycles for all $m \leq 7$ but it cannot even detect them for $m > 7$.
\end{prop}
\begin{hproof}
	Only the idea behind triangle and 4-cycle counting is outlined here to give an intuition for why \cref{prop:related:wl2-cycle-count} holds.
	Note that the 2-\acs{wl} neighborhood of each edge $(v, w) \in \mathcal{E}_G$ includes all possible paths $(v, u, w)$ of length $2$, i.e.\ all possible triangles.
	By \cref{defn:related:wlk-refine} there must thus be a color subset $C^{\triangle}_j \subseteq \mathcal{C}$ representing that an edge is involved in exactly $j$ triangles after one refinement step.
	2-\acs{wl} then trivially counts triangles by setting $g(\mathit{dist}_\chi) = \frac{1}{3} \sum_j j \sum_{c \in C^{\triangle}_j} \mathit{dist}_\chi(c)$ to satisfy \cref{defn:related:wl-compute}.
	Analogously for 4-cycles, let $C^{\square}_j \subseteq \mathcal{C}$ be the colors indicating a non-edge $(v, w) \notin \mathcal{E}_G$ with $j$ common vertex neighbors $u \in \Gamma_G(v) \cap \Gamma_G(w)$.
	The number of 4-cycles is then determined by the colors of the diagonals through them via $g(\mathit{dist}_\chi) = \frac{1}{2} \sum_{j \geq 2} \binom{j}{2} \sum_{c \in C^{\square}_j} \mathit{dist}_\chi(c)$.
	Using a similar but more involved combinatorial argument requiring multiple color refinement steps, 5-, 6- \& 7-cycle counting can be shown.
\end{hproof}
As we just saw, 2-\acs{wl} is significantly more powerful than 1-\acs{wl}.
Though, by \cref{prop:related:wl-upper}, there are of course still 2-\acs{wl} indistinguishable graphs;
among others those are the strongly regular graphs $\mathit{srg}_{n, d, \lambda, \mu}(G)$.
\begin{defn}
	$\begin{aligned}[t]
		\mathit{srg}_{n, d, \lambda, \mu}(G) \iff\ & {|\mathcal{V}_G|} = n \quad\land\quad \forall v \in \mathcal{V}_G: {|\Gamma_G(v)|} = d \\
		&\land \forall (v, w) \in \mathcal{E}_G: {|\Gamma_G(v) \cap \Gamma_G(w)|} = \lambda \\
		&\land \forall (v, w) \in \mathcal{V}_G^2 \setminus \mathcal{E}_G: {|\Gamma_G(v) \cap \Gamma_G(w)|} = \mu
	\end{aligned}$
\end{defn}
Generally this restriction of 2-\acs{wl} is not an issue since strongly regular graphs do not appear in typical \ac{gcr} datasets.
By going to $k = 3$, even some strongly regular graphs as well as all planar graphs can be distinguished though~\cite{Kiefer2017}.
Apart from that there are currently few results regarding the classes of distinguishable graphs and computable functions for even higher \ac{wl}-dimensions.

@@

\paragraph{A 1-\acs{wl} bound on \ac{gnn} power}
In the next step we will take a closer look at the relation between \acp{gcnn} that use vertex neighborhood convolution and 1-\acs{wl} by comparing their respective discriminative power.
\begin{prop}[see {\citet[lemma 2 and theorem 3]{Xu2018}} for the proof]\label{prop:related:gcnn-wl1-limit}
	The discriminative power of a \ac{gcnn} that convolves vertex feature vectors via a convolution operator of the form $Z_G^{(t)}[v] = h^{(t)}\left(Z_G^{(t-1)}[v], \ldblbrace Z_G^{(t-1)}[u]\, |\, u \in \Gamma_G(v) \rdblbrace\right)$ is upper-bounded by that of 1-\acs{wl}, where $h^{(t)}: {\left(\mathbb{R}^{d^{(t-1)}}\right)}^{*} \to \mathbb{R}^{d^{(t)}}$ is an arbitrary vertex neighborhood hashing function.
	Moreover, iff.\ $h^{(t)}$ is injective, the \ac{gcnn} has the same discriminative power as 1-\acs{wl}.
\end{prop}
By \cref{prop:related:gcnn-wl1-limit} the \ac{gcn} architecture is strictly less powerful than 1-\acs{wl} due to the fact that $h_{\mathrm{\acs*{gcn}}}$ is not an injective hash function.
To overcome this limitation \citet{Xu2018} proposed an alternative vertex neighborhood convolution architecture called \ac{gin}:
\begin{align}
	Z_G^{(t)}[v] \coloneqq \mathrm{\acs*{mlp}}^{(t)}\left( Z_G^{(t-1)}[v] + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} Z_G^{(t-1)}[u] \right) \label{eq:related:gin-layer}
\end{align}
By leaving out the normalization factors $\mu(v)$ and by using a \ac{mlp} instead of the single fully-connected layer $\sigma \circ \Theta^{(t)}$ the resulting hashing function $h_{\mathrm{\acs*{gin}}}(z_1, \dots, z_m) = \mathrm{\acs*{mlp}}^{(t)}(\sum_{j=1}^{m} z_j)$ becomes injective\footnote{
	Under the assumption that the set of possible input feature matrices $X \in \mathbb{R}^{n \times d^{(0)}}$ is countable.
	This is reasonable since real-world \ac{gcr} domains typically only have a finite set of possible vertex feature vectors, e.g\ one-hot label encodings.
}, i.e.\ it assigns a unique color vector to each multiset of color vectors.
By \cref{prop:related:gcnn-wl1-limit} this implies that \ac{gin} has the same discriminative power as 1-\ac{wl}.
As expected, \ac{gin} therefore fits training data better than non-injective \acp{gcnn} like \ac{gcn}.
Interestingly \ac{gin} additionally seems to generalize better on test data than non-injective methods~\cite{Xu2018}\cite{Errica2020}.
It is not yet fully understood why this is the case.
